---
title: "Extracting field boundaries from NASS CropSpace CDL data"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{css}
p {
  font-size: 13pt;
}
```

This tutorial shows how to extract field boundaries from CropSpace CDL data available at NASS. This procedure involves several steps:

1. Fecthing raster data of crop coverage from NASS API.
2. Selecting crops of interest and masking other crops
3. Detecting boundaries and mask them out
4. Detecting cumps 
5. Vectorize clumps
6. Remove polygons by area

## Step 1

For this example we are going to use the data from Champaign County (IL) year 2018. According to the Federal Information Processing System (FIPS), the code for the state is `17` and for the county `019`. Then, the full code is: `17019`.

```{r}
FIPS <- 17019
year <- 2018
base_url <- "https://nassgeodata.gmu.edu/webservice/nass_data_cache/byfips/CDL"
req_url <- paste0(base_url, "_", year, "_", FIPS, ".tif")
req_rul
```


Once we have the request url we can download the `tif` image into R using `raster::raster()`.

```{r}
library(raster)
cdl_rst <- raster(req_url)
cdl_rst
```

For visualizing the results we can use the `tmap` package

```{r}
library(tmap)
tm_shape(cdl_rst) +
  tm_raster("CDL_2018_17019.tif", drop.levels = T, showNA = F, legend.show = F)
```


```{r, include = F, eval =F}
crp_rst <- raster(xmn=640000, xmx=650000, ymn=1899000, ymx=1910000,
                  res = 900, crs = crs(cdl_rst))

cdl_rst2 <- crop(cdl_rst, crp_rst)
```


Based on the plot above, the most common category are green and yellow. To find out which crop are they we need the values attribute table (VAT) info which is stored in the `generic_cdl_attributes.tif.var.dbf.zip`.

```{r}
vat_url <- "https://www.nass.usda.gov/Research_and_Science/Cropland/docs/generic_cdl_attributes.tif.vat.dbf.zip"
tmp_fn <- tempfile()
download.file(vat_url, destfile = tmp_fn)
file.exists(tmp_fn)
```

As it is a ZIP file we can chechk its content using `unzip()` with `list = T`.

```{r}
unzip(zipfile = tmp_fn, list = T)
```

To read the information we need to extract the zipfile in a temporary directory and then read the dbf database `ArcGIS10.3.1_2019_30m_cdls.img.vat.dbf`.

```{r}
# Extract the files
tmp_dir <- tempdir()
unzip(zipfile = tmp_fn, files = "ESRI_attribute_files/ArcGIS10.3.1_2019_30m_cdls.img.vat.dbf", 
      exdir = tmp_dir)

# Read
library(foreign)
cdl_vat <- read.dbf(paste0(tmp_dir, "/ESRI_attribute_files/ArcGIS10.3.1_2019_30m_cdls.img.vat.dbf"))
head(cdl_vat, n = 15)
```

## Step 2

To see what are the most common values we need to tabulate the values observed in the CDL data raster and join those values with their category names

```{r}
library(dplyr)
tbl <- data.frame(table(values(cdl_rst), dnn = "VALUE"), stringsAsFactors = F)
cdl_vat$VALUE <- factor(cdl_vat$VALUE)
tbl <- left_join(tbl, cdl_vat, by = "VALUE") %>% 
  arrange(desc(Freq))
tbl
```

We cab see that the most common crops within Champaign County are `Corn` and `Soybeans`. We are going to use this information for masking out those areas which are not of interest.

```{r}
# Create mask
cdl_msk <- cdl_rst
cdl_msk[!cdl_msk %in% c(1, 5)] <- NA 
cdl_rst <- mask(cdl_rst, mask = cdl_msk)
tm_shape(cdl_rst) +
  tm_raster("CDL_2018_17019", drop.levels = T, showNA = F, legend.show = F)
```
## Step 3

We have identified all the agricultural fields planted with corn and soybeans. Next step is to detect boundaries.


```{r}
# Detect boundaries
cdl_bnd <- boundaries(cdl_rst, classes = T, type = "inner", asNA = T)
tm_shape(cdl_bnd) +
  tm_raster("layer", drop.levels = T, showNA = F, legend.show = F)
```

In the map above, colored pixels represents arreas considered borders. We can use this information to derive a inverse mask

```{r}
# Mask out boundaries
cdl_rst <- mask(cdl_rst, mask = cdl_bnd, inverse = T)
tm_shape(cdl_rst) +
  tm_raster("CDL_2018_17019", drop.levels = T, showNA = F, legend.show = F)
```
## Detect clumps

Having the fields identified and masked them out from non-interest areas and boundaries, we can proceed to detect clumps which area sets of neighbouring pixeles.

```{r}
system.time(
  cdl_clmp <- clump(cdl_rst, directions = 4, gaps = F)
)
tm_shape(cdl_clmp) +
  tm_raster("clumps", drop.levels = T, showNA = F, legend.show = F)
```

These clumps represents groups of pixels which can be considered as a independent field.

## Vectorized clumps

Using the previous map we can turn the raster dataset into multipolygon data.

```{r}
beginCluster(10)

# Vectorize those clumps
system.time(
  cdl_vct <- rasterToPolygons(cdl_clmp, dissolve = T)
)
tm_shape(cdl_clmp) +
  tm_polygons("clumps", drop.levels = T, showNA = F, legend.show = F)
```

## Refine map

```{r}
# convert to spatial feature and remove holes
cdl_vct <- st_as_sf(cdl_vct)
cdl_vct <- nngeo::st_remove_holes(cdl_vct)
tm_shape(cdl_clmp) +
  tm_polygons("clumps", drop.levels = T, showNA = F, legend.show = F)
```

```{r}
# Remove small fields
cdl_vct <- mutate(cdl_vct, area = st_area(cdl_vct)) %>% 
  filter(area > as_units(100000, "m^2"))

tm_shape(cdl_vct) +
  tm_polygons("clumps", drop.levels = T, showNA = F, legend.show = F)

```

## Write data

```{r}
# Write data
write_sf(cdl_vct, "cdl_tmp.gpkg")
```